Type Error Telemetry at Scale

### 2023-09-24 programming23/paper.tex
sec 3 relist telemetry fields, misc other fixes (captions)

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FILL}{\textbf{FILL}}
\newcommand{\dotscale}[1]{\scalebox{0.72}{#1}}
\newcommand{\wideas}[2]{\makebox[\widthof{#2}][l]{#1}}
\newcommand{\twoline}[2]{\parbox[s]{1.4cm}{\flushleft#1\newline#2}}
\newcommand{\chkYes}{\dotscale{\CIRCLE}}
\newcommand{\chkMaybe}{\wideas{\dotscale{\Circle}}{\chkYes}}
\newcommand{\chkNo}{\wideas{}{\chkYes}}
\newcommand{\pct}[1]{\SI{#1}{\percent}}
\newcommand{\percentile}[1]{{#1}th~percentile}
\newcommand{\modefont}[1]{\texttt{#1}}
\newcommand{\mnocheck}{\modefont{nocheck}}
\newcommand{\mnonstrict}{\modefont{nonstrict}}
\newcommand{\mstrict}{\modefont{strict}}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\nK}[1]{#1\textsc{k}}
\newcommand{\stddev}[1]{[\nK{#1}]}
\newcommand{\gcell}[1]{\cellcolor{green!20}#1}
\newcommand{\ycell}[1]{\cellcolor{yellow!18}#1}
\newcommand{\ocell}[1]{\cellcolor{orange!29}#1}
\newcommand{\rcell}[1]{\cellcolor{red!30}$\!\!$#1$\!\!$}
\newcommand{\gbox}[1]{\colorbox{green!20}{#1}}
\newcommand{\ybox}[1]{\colorbox{yellow!18}{#1\vphantom{Ap}}}
\newcommand{\obox}[1]{\colorbox{orange!29}{#1\vphantom{Ap}}}
\newcommand{\rbox}[1]{\colorbox{red!30}{#1\vphantom{Ap}}}
\newcommand{\tefsfont}[1]{\textsc{#1}}
\newcommand{\tekey}{\tefsfont{te}}
\newcommand{\fskey}{\tefsfont{fs}}
\newcommand{\QALAN}{\FILL{} Alan:}

@@

In this paper, we investigate methods for measuring the effectiveness
of the {Luau} type system in the development of {Roblox} scripts.
We want to collect measurements at a large scale, with thousands
of participants, and we must do so in a way that protects creators' privacy.
In comparison to prior work~(\cref{s:related}), which
with few exceptions~\cite{zhlbr-cc-2020,zhlbr-oopsla-2020,hlzbr-ecoop-2021} is either small in scale
or collects personally identifiable information~(PII),
we performed a large-scale study using pseudonymized \emph{telemetry}.

@@

The telemetry data we analyzed does not contain any PII.
It has no source code;
 no source locations;
no error messages (which may contain source code);
no error messages;
no record of the creator's identity, locale, or IP address;
and no information about what creation the data came from.
Telemetry records are correlated by session, using a pseudonymized
session identifier.

@@

\section{{Roblox} Context}
% https://create.roblox.com/docs/scripting/luau

@@

  \caption{{Roblox Studio 3D creation} tools (left) and IDE (right)}
  \label{fig:roblox-studio}
\end{figure}

@@

{Roblox Studio} is a workbench that
combines {3D creation} tools and an Integrated
Developer Environment (IDE), as seen in \Cref{fig:roblox-studio}.
The IDE includes an optional ``Script Analysis'' widget, which
reports syntax errors, type errors, and problems identified by
lint tools. The script editor also (optionally) highlights
the location in code where reported errors occur.
To opt in to type error reporting, creators set a \emph{mode}
for each script, which is one of:
\begin{itemize}
  \item \mnocheck{}: only syntax errors are reported,
  \item \mnonstrict{}: all syntax errors, and a subset of ``high confidence'' type errors, are reported, or
  \item \mstrict{}: all syntax errors and type errors, are reported.
\end{itemize}
As an example of nonstrict mode, the following program only reports one error:
\begin{verbatim}
  --!nonstrict
  local x = { p = 5, q = nil }
  if condition then x.q = 7 end
  local y = x.p + x.q --> no type error
  local z = x.r       --> "Key 'r' not found in table 'x'"
\end{verbatim}
but in strict mode it reports two:
\begin{verbatim}
  --!strict
  local x = { p = 5, q = nil }
  if condition then x.q = 7 end
  local y = x.p + x.q --> "Type 'nil' could not be converted into 'number'"
  local z = x.r       --> "Key 'r' not found in table 'x'"
\end{verbatim}
In cases like this, where it is undecidable whether there will be a run-time error,
strict mode errs on the side of reporting an error, and nonstrict mode errs on
the side of suppressing the error.

@@

\begin{table}
  \caption{Selected Error Labels}
  % https://github.com/Roblox/luau/blob/master/Analysis/include/Luau/Error.h
  %% from type analysis, but they're not all really type errors
  %% TODO review selection after analyzing the full data
  \label{t:type-error-labels}

@@

The telemetry for a Roblox Studio session is gathered as a series of
telemtry records, all with the same psedonynimized session
identifier. This allows us to correlate telemetry across a single
session, but not between sessions.

In order to avoid swamping our telemetry servers, we randomly throttle
the generation of telemetry records:
\begin{itemize}
  \item
    1\% of Roblox Studio sessions were enrolled in generating telemetry, and
  \item
    0.5\% of runs of the typechecker (approximately 1 in 200 keystrokes)
      in an enrolled session generated a telemetry record.
\end{itemize}
In order to find how file-switching correlated with type errors, we additionally have:
\begin{itemize}
  \item
    every run of the typechecker in an enrolled session which switched file generated a telemetry record.
\end{itemize}
Each telemetry record contains seven kinds of data.
Refer to~\cref{f:tdata} for an illustration:
\begin{enumerate}
  \item
    ID: a psedonynimized identifier for the current session
  \item
    Time: one timestamp from the client and one from the server.
    %% timestamp explanation \cref{s:data-cleaning}
  \item
    Mode: type analysis mode of the current file, which is either \mnocheck{},
    \mnonstrict{}, or \mstrict{}.
%% Reason & Size Info & Overall Counts & Edit Range Counts

  \item reason for sending: (random selection, or file switch),
  \item size of codebase: number of files, and number of lines of code type checked,
  \item size of edit range
  \item overall type errors: for previous and current states, we have the total number, the nunber in the current file, and number in the edit range,
  \item specific errors in the edit range: for the previous and current states, for each type error code, the number type errors in the edit range,
    \FILL{} explain previous/current in detail, explain that this data takes up most of the telemetry record
  \item forced strict errors: specific errors for the
    edit range computed as if the user had strict mode enabled.
  \item TooComplexErrors: project-wide count for this one specific error type
     \cref{s:code-too-complex}
\end{enumerate}
To track the edit range, we record a start and end position, which we
update appropriately on every edit. This can result in very large edit
ranges, for example, if the user edits at the beginning and end of the
file.
The main benefit of this strategy is that it ensures constant-size telemetry records.

@@

  \caption{Session size in seconds and in number of records}
  \label{f:session-size}
\end{figure}

@@

\begin{figure}[t]
  \begin{tabular}[t]{ll} \\
    \begin{tabular}[t]{r@{~~}r@{~}l}
      595,137 & type errors \\
      289,698 & in current module & [\pct{48.68}] \\
       30,924 & in edit range & [\pct{5.20}]
    \end{tabular}
    \begin{tabular}[t]{r@{~~}r@{~}l}
      72,235,735 & {forced strict errors} \\
      37,027,281 & in current module & [\pct{51.26}] \\
       1,111,178 & in edit range & [\pct{1.54}]
    \end{tabular}
  \end{tabular}
  \caption{Type errors and forced strict errors across all telemetry records}
  \label{f:count-analysis-errors}
\end{figure}

@@

\begin{figure}[t]\centering
  \begin{subfigure}[t]{\columnwidth}
    \begin{tabular}[t]{ll}
      %% FILL bar chart? rectangle?
      \begin{tabular}[t]{r@{~~}l@{~}r}
         1,341,348 & \mnocheck{}          & [\pct{89.14}] \\
           156,883 & \mnonstrict{}        & [\pct{10.43}] \\
             6,505 & \mstrict{}           & [\pct{ 0.43}]
      \end{tabular}
      \begin{tabular}[t]{r@{~~}l@{~~}r}
      \end{tabular}
    \end{tabular}
    \caption{Analysis modes across the 1,504,736 records}
    \label{f:total-records}
  \end{subfigure}

  \begin{subfigure}[t]{\columnwidth}
    \begin{tabular}[t]{ll} \\
      \begin{tabular}[t]{r@{~~}r@{~}r}
        313,509 & \mnocheck{}   & [\pct{90.19}] \\
         32,902 & \mnonstrict{} & [\pct{ 9.47}] \\
            545 & \mstrict{}    & [\pct{ 0.16}] \\
            642 & mixed-mode    & [\pct{ 0.18}]
      \end{tabular}
      \begin{tabular}[t]{l@{~~}ll}
        \zerowidth{Among the mixed-mode sessions:} \\
        & 341 & contain a mode upgrade \\
        & 320 & contain a mode downgrade \\
        & 512 & contain modules with different modes
      \end{tabular}
    \end{tabular}
    \caption{Analysis mode(s) for the 347,598 sessions}
    \label{f:total-sessions}
  \end{subfigure}

  \caption{Overview of type analysis modes}
  \label{f:dataset-overview}
\end{figure}

@@

\Cref{t:type-error-count} reports the percent of
type errors that overlap with the edit range.
This table is based on the same data as the previous table~(\cref{t:type-error-survival}),
but focuses on the overall popularity of type errors
rather than their relation to edits.
The left column in each row names one kind of type error
that overlaps with an edit region.
The right column shows what percentage of all overlapping
errors the current error took up.

@@

Fortunately, the data rarely contains too-complex errors.
None of these errors appear in edit regions~(\cref{t:type-error-count}).
For the specific case of \code{CodeTooComplex}, our telemetry tracked project-wide
counts and found only 26 of these errors, which were spread across eleven
records in three sessions.
It is entirely possible that all these errors came from one creator who
deliberately pushed the limits of type analysis.
Whatever the reason, the overall number is low enough to suggest that the
internal limits are not posing barriers to typechecker adoption.

@@

  \mnocheck{}
  \includegraphics[width=\columnwidth]{img/error-count-nocheck-row--te-density-diff.pdf}
  \medskip
  \mnonstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-nonstrict-row--te-density-diff.pdf}
  \medskip
  \mstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-strict-row--te-density-diff.pdf}
  \caption{Changes over time (seconds) to type error density (total errors / lines in project)}
  \label{f:error-density}
\end{figure}

@@

\begin{figure}[t]\centering
  %% TODO
  %% - thin bars
  %% - colors for good , bad (disagree) , neutral
  \begin{tabular}{lll}
    \mnocheck{} & \mnonstrict{} & \mstrict{} \\
    \includegraphics[width=0.3\columnwidth]{img/compass-nocheck.pdf}
    &
    \includegraphics[width=0.3\columnwidth]{img/compass-nonstrict.pdf}
    &
    \includegraphics[width=0.3\columnwidth]{img/compass-strict.pdf}
  \end{tabular}
  \caption{FILL Are there fewer type errors than forcedstrict errors? Expect true always, but that is not the case.}
  \label{f:tefs-compass}
\end{figure}

@@

  \caption{Example sessions}
  \label{f:indy-session}
\end{figure}

@@

Telemetry cannot replace user studies, as there is no way to measure
creator sentiment, but provide complementary data at scale.

@@

\begin{table}[t]\centering
  \caption{Comparison of telemetry designs}
  \label{t:telemetry-design}

  \begin{tabular}{l@{~~}c@{~~}c@{~~}c@{~~}c@{~~}c}
    ~             & Event Counts & Random  & Timestamps & Error Msgs. & Source Code \\\midrule
    Roblox Studio & \chkYes      & \chkYes & \chkYes    & \chkNo        & \chkNo    \\
    Transparent   & \chkYes      & \chkNo  & \chkNo     & \chkNo        & \chkNo    \\
    Classic       & \chkYes      & \chkNo  & \chkYes    & \chkYes       & \chkYes   \\
%% orthogonal %%    & Diff.Priv
%% orthogonal %%    & CBI
  \end{tabular}
\end{table}

>>>>

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FILL}{\textbf{FILL}}
\newcommand{\dotscale}[1]{\scalebox{0.72}{#1}}
\newcommand{\wideas}[2]{\makebox[\widthof{#2}][l]{#1}}
\newcommand{\twoline}[2]{\parbox[s]{1.4cm}{\flushleft#1\newline#2}}
\newcommand{\chkYes}{\dotscale{\CIRCLE}}
\newcommand{\chkMaybe}{\wideas{\dotscale{\Circle}}{\chkYes}}
\newcommand{\chkNo}{\wideas{}{\chkYes}}
\newcommand{\pct}[1]{\SI{#1}{\percent}}
\newcommand{\percentile}[1]{{#1}th~percentile}
\newcommand{\modefont}[1]{\texttt{#1}}
\newcommand{\mnocheck}{\modefont{nocheck}}
\newcommand{\mnonstrict}{\modefont{nonstrict}}
\newcommand{\mstrict}{\modefont{strict}}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\nK}[1]{#1\textsc{k}}
\newcommand{\stddev}[1]{[\nK{#1}]}
\newcommand{\gcell}[1]{\cellcolor{green!20}#1}
\newcommand{\ycell}[1]{\cellcolor{yellow!18}#1}
\newcommand{\ocell}[1]{\cellcolor{orange!29}#1}
\newcommand{\rcell}[1]{\cellcolor{red!30}$\!\!$#1$\!\!$}
\newcommand{\gbox}[1]{\colorbox{green!20}{#1}}
\newcommand{\ybox}[1]{\colorbox{yellow!18}{#1\vphantom{Ap}}}
\newcommand{\obox}[1]{\colorbox{orange!29}{#1\vphantom{Ap}}}
\newcommand{\rbox}[1]{\colorbox{red!30}{#1\vphantom{Ap}}}
\newcommand{\tefsfont}[1]{\textsc{#1}}
\newcommand{\tekey}{\tefsfont{te}}
\newcommand{\fskey}{\tefsfont{fs}}
\newcommand{\QALAN}{\FILL{} Alan:}
\newcommand{\panon}{pseudonymized}

@@

In this paper, we investigate methods for measuring the effectiveness
of the {Luau} type system in the development of {Roblox} scripts.
We want to collect measurements at a large scale, with thousands
of participants, and we must do so in a way that protects creators' privacy.
In comparison to prior work~(\cref{s:related}), which
with few exceptions~\cite{zhlbr-cc-2020,zhlbr-oopsla-2020,hlzbr-ecoop-2021} is either small in scale
or collects personally identifiable information~(PII),
we performed a large-scale study using \panon{} \emph{telemetry}.

@@

The telemetry data we analyzed does not contain any PII.
It has no source code;
 no source locations;
no error messages (which may contain source code);
no error messages;
no record of the creator's identity, locale, or IP address;
and no information about what creation the data came from.
Telemetry records are correlated by session, using a \panon{}
session identifier.

@@

\section{{Roblox} Context}
\label{s:context}
% https://create.roblox.com/docs/scripting/luau

@@

  \caption{{Roblox Studio 3D creation} tools (left) and IDE (right).}
  \label{fig:roblox-studio}
\end{figure}

@@

{Roblox Studio} is a workbench that
combines {3D creation} tools and an Integrated
Developer Environment (IDE), as seen in \Cref{fig:roblox-studio}.
The IDE includes an optional \emph{Script Analysis} widget that
reports syntax errors, type errors, and problems identified by
lint tools. The script editor also (optionally) highlights
the location in code where reported errors occur.
Creators have three \emph{analysis modes} (modes) to choose from
for every script they write:
\begin{itemize}
  \item \mnocheck{}: report only syntax errors,
  \item \mnonstrict{}: report syntax errors and a subset of ``high confidence'' type errors, or
  \item \mstrict{}: report syntax errors and all type errors.
\end{itemize}
Each run of type analysis can report several errors.
There is no guarantee that creators read every error.
In fact, creators who close the Script Analysis widget
will see highlights in their code but no further details.

As an example of nonstrict mode, the following program only reports one error:
\begin{verbatim}
  --!nonstrict
  local x = { p = 5, q = nil }
  if condition then x.q = 7 end
  local y = x.p + x.q --> no type error
  local z = x.r       --> "Key 'r' not found in table 'x'"
\end{verbatim}
but in strict mode it reports two:
\begin{verbatim}
  --!strict
  local x = { p = 5, q = nil }
  if condition then x.q = 7 end
  local y = x.p + x.q --> "Type 'nil' could not be converted into 'number'"
  local z = x.r       --> "Key 'r' not found in table 'x'"
\end{verbatim}
In cases like this, where it is undecidable whether there will be a run-time error,
strict mode errs on the side of reporting an error, and nonstrict mode errs on
the side of suppressing the error.

@@

\begin{table}
  \caption{Sample of type errors.}
  % https://github.com/Roblox/luau/blob/master/Analysis/include/Luau/Error.h
  %% from type analysis, but they're not all really type errors
  \label{t:type-error-labels}

  %% TODO review selection after analyzing the full data

@@

The telemetry for a Roblox Studio session is gathered as a series of
telemtry records, all with the same \panon{} session
identifier. This allows us to correlate telemetry across a single
session, but not between sessions.

In order to avoid swamping our telemetry servers, we randomly throttle
the generation of telemetry records:
\begin{itemize}
  \item
    1\% of Roblox Studio sessions were enrolled in generating telemetry, and
  \item
    0.5\% of runs of the typechecker (approximately 1 in 200 keystrokes)
      in an enrolled session generated a telemetry record.
\end{itemize}
In order to find how file-switching correlated with type errors, we additionally have:
\begin{itemize}
  \item
    every run of the typechecker in an enrolled session which switched file generated a telemetry record.
\end{itemize}

To track the edit range, we record a start and end position, which we
update appropriately on every edit. This can result in very large edit
ranges, for example, if the user edits at the beginning and end of the
file.
The main benefit of this strategy is that it ensures constant-size telemetry records.

Each telemetry record contains seven kinds of data.
Refer to~\cref{f:tdata} for an illustration:
\begin{enumerate}
  \item
    ID: a \panon{} identifier for the current session
  \item
    Time: one timestamp from the client and one from the server.
    %% timestamp explanation \cref{s:data-cleaning}
  \item
    Mode: type analysis mode of the current file, which is either \mnocheck{},
    \mnonstrict{}, or \mstrict{}.
  \item
    Reason: a flag that explains whether this record was sent due to a
    randomly-selected keystroke or a module switch.
  \item
    Size Info: number of lines in the codebase and number of lines in the edit range.
  \item
    Overall Counts: summarizes type errors and forced strict errors during the
    last two invocations of type analysis, and additionally reports the number
    of times type analysis hit an internal limit on problem
    size~(see~\cref{s:code-too-complex} for details).
    For each invocation and each kind of
    analysis error, there are three summary counts:
    %% grand total = 12
    \begin{itemize}
      \item total number of errors,
      \item errors in the current module, and
      \item errors in the current edit range.
    \end{itemize}
  \item
    Edit Range Counts: these list specific errors that arose in the last two
    type analyses and overlap with the current edit range.
    Since there are 35 possible errors~(\cref{s:context}), there can be up to 70 counts
    in a telemetry record.
    Exactly how to interpret this data depends on which analysis it came from:
    \begin{itemize}
      \item
        Errors in the latest type analysis are currently visible to the
        creator.
        These may have been introduced by the changes in the edit range.
      \item
        Errors from the previous type analysis were visible before the creator
        made the latest edits.
        These may have motivated some of the changes in the edit range.
    \end{itemize}
\end{enumerate}

\paragraph{Reflections}
This kind of telemetry data is coarse-grained.
For instance, it does not let us distinguish between edits that ignore a type
error from edits that remove one error while introducing another.
We acknowledge this and other threats in~\cref{s:threats}.
The main advantages of the telemetry design are the small size of each record
and complete the lack of PII.
Furthermore, despite the limitations, Studio's telemetry supports a variety of
interesting conclusions that we showcase in the next section.

@@

  \caption{Session size in seconds and in number of records.}
  \label{f:session-size}
\end{figure}

@@

\begin{figure}[t]
  \begin{tabular}[t]{ll} \\
    \begin{tabular}[t]{r@{~~}r@{~}l}
      595,137 & type errors \\
      289,698 & in current module & [\pct{48.68}] \\
       30,924 & in edit range & [\pct{5.20}]
    \end{tabular}
    \begin{tabular}[t]{r@{~~}r@{~}l}
      72,235,735 & {forced strict errors} \\
      37,027,281 & in current module & [\pct{51.26}] \\
       1,111,178 & in edit range & [\pct{1.54}]
    \end{tabular}
  \end{tabular}
  \caption{Type errors and forced strict errors across all telemetry records.}
  \label{f:count-analysis-errors}
\end{figure}

@@

\begin{figure}[t]\centering
  \begin{subfigure}[t]{\columnwidth}
    \begin{tabular}[t]{ll}
      %% FILL bar chart? rectangle?
      \begin{tabular}[t]{r@{~~}l@{~}r}
         1,341,348 & \mnocheck{}          & [\pct{89.14}] \\
           156,883 & \mnonstrict{}        & [\pct{10.43}] \\
             6,505 & \mstrict{}           & [\pct{ 0.43}]
      \end{tabular}
      \begin{tabular}[t]{r@{~~}l@{~~}r}
      \end{tabular}
    \end{tabular}
    \caption{Analysis modes across the 1,504,736 records.}
    \label{f:total-records}
  \end{subfigure}

  \begin{subfigure}[t]{\columnwidth}
    \begin{tabular}[t]{ll} \\
      \begin{tabular}[t]{r@{~~}r@{~}r}
        313,509 & \mnocheck{}   & [\pct{90.19}] \\
         32,902 & \mnonstrict{} & [\pct{ 9.47}] \\
            545 & \mstrict{}    & [\pct{ 0.16}] \\
            642 & mixed-mode    & [\pct{ 0.18}]
      \end{tabular}
      \begin{tabular}[t]{l@{~~}ll}
        \zerowidth{Among the mixed-mode sessions:} \\
        & 341 & contain a mode upgrade \\
        & 320 & contain a mode downgrade \\
        & 512 & contain modules with different modes
      \end{tabular}
    \end{tabular}
    \caption{Analysis mode(s) for the 347,598 sessions.}
    \label{f:total-sessions}
  \end{subfigure}

  \caption{Overview of type analysis modes.}
  \label{f:dataset-overview}
\end{figure}

@@

\Cref{t:type-error-count} reports the percent of
type errors that overlap with the edit range.
This table is based on the same data as the previous table~(\cref{t:type-error-survival}),
but focuses on the overall popularity of type errors
rather than their relation to edits.
The left column in each row names one kind of type error
that overlaps with an edit range.
The right column shows what percentage of all overlapping
errors the current error took up.

@@

Fortunately, the data rarely contains too-complex errors.
None of these errors appear in edit ranges~(\cref{t:type-error-count}).
For the specific case of \code{CodeTooComplex}, our telemetry tracked project-wide
counts and found only 26 of these errors, which were spread across eleven
records in three sessions.
It is entirely possible that all these errors came from one creator who
deliberately pushed the limits of type analysis.
Whatever the reason, the overall number is low enough to suggest that the
internal limits are not posing barriers to typechecker adoption.

@@

  \mnocheck{}
  \includegraphics[width=\columnwidth]{img/error-count-nocheck-row--te-density-diff.pdf}
  \medskip
  \mnonstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-nonstrict-row--te-density-diff.pdf}
  \medskip
  \mstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-strict-row--te-density-diff.pdf}
  \caption{Changes over time (seconds) to type error density (total errors / lines in project).}
  \label{f:error-density}
\end{figure}

@@

\begin{figure}[t]\centering
  %% TODO
  %% - thin bars
  %% - colors for good , bad (disagree) , neutral
  \begin{tabular}{lll}
    \mnocheck{} & \mnonstrict{} & \mstrict{} \\
    \includegraphics[width=0.3\columnwidth]{img/compass-nocheck.pdf}
    &
    \includegraphics[width=0.3\columnwidth]{img/compass-nonstrict.pdf}
    &
    \includegraphics[width=0.3\columnwidth]{img/compass-strict.pdf}
  \end{tabular}
  \caption{Comparing the change in type errors to the change in forced strict errors.}
  \label{f:tefs-compass}
\end{figure}

@@

  \caption{Two example sessions.}
  \label{f:indy-session}
\end{figure}

@@

We are sampling sessions. Could miss interesting ones.

Sampling by keystroke means different things for different creators.
Depends how they type.
Scripted interactions that generate a large number of keystrokes per second
can swamp our data and are hard to distinguish from genuine interactions.
Copy paste might just look like a bot.
\QALAN{} I think you said a paste counts as 1 keystroke; is that right?

We have no knowledge whatsoever about interesting events like save run publish.
Publish events are private.

Changes to errors in the edit range are hard to follow.
Could add one error but remove another of the same kind,
then we can't distinguish from ignoring the error.

Edit ranges are coarse, merely the top and bottom edit.
No idea what happened in between.
Long time between keystroke events makes this even worse.

Telemetry cannot replace user studies, as there is no way to measure
creator sentiment, but provide complementary data at scale.
No idea about intent is a big problem, we don't know which error the creator wanted to fix.

Type analysis reports many errors at once.
No way to know what the creator is looking at.
Not even sure they have the analysis widget open.

Internal validity?

External validity?

Construct validity?

Conclusion validity

Ecological validity is high.

@@

\begin{table}[t]\centering
  \caption{Comparison of telemetry designs. Two orthogonal concerns are when to collect data and whether to add noise to it.}
  \label{t:telemetry-design}

  \begin{tabular}{l@{~~}c@{~~}c@{~~}c@{~~}c@{~~}c}
    ~             & Event Counts & Timestamps & Error Msgs. & Source Code \\\midrule
    Roblox Studio & \chkYes      & \chkYes    & \chkNo        & \chkNo    \\
    Transparent   & \chkYes      & \chkNo     & \chkNo        & \chkNo    \\
    Classic       & \chkYes      & \chkYes    & \chkYes       & \chkYes   \\
  \end{tabular}
\end{table}
